### 2018-06-21 algorithm.tex
labor adjustment cost writeup

\subsection{Time series}
Given the autoregressive nature of productivity, the vector of $A_{jnt}$s is a state variable. The problem is drastically simplified by the fact that no decision variable affects the state. The representative agent solves
\[
\max_{\{L_{njt}, L_{njt}^*\}}
E
\sum_{t=0}^{\infty} 
	\delta^{t} 
	\left[
	\ln \frac 
		{\sum_j w_{njt}(\mathbf A_{t}, \mathbf L_{t})L_{njt}}
		{P_{nt}(\mathbf A_{t}, \mathbf L_{t})}
	-\frac \varrho 2
		\sum_{j=1}^J
		\left(
			\frac {L_{njt}} {L_{nt}}
			- 
			\frac {L_{njt}^*} {L_{nt}}
		\right)^2
	\right]
\]
subject to $\sum_j L_{njt}=\sum_j L_{njt}^*=L_{nt}$ and $w$ and $P$ being determined in equilibrium. The starred labor allocation is determined before shock realizations are known.

The corresponding Bellman equation is
\begin{equation}
	V_n^N(\mathbf A_{t-1}) = 
	\max_{L_{njt}^*}
	E_{A} V_n^D(\mathbf A_{t}, \mathbf L_{t}^*)
\end{equation}
for the ``night'' period before productivity shocks have been realized. For the ``day'' period,
\begin{equation}
	V_n^D(\mathbf A_{t}, \mathbf L_{t}^*) = 
	\max_{\{L_{njt}\}}
		\ln \frac 
			{\sum_j w_{njt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)L_{njt}}
			{P_{nt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)}
		-\frac \varrho 2
			\sum_{j=1}^J
			\left(
				\frac {L_{njt}} {L_{nt}}
				- 
				\frac {L_{njt}^*} {L_{nt}}
			\right)^2
	+ \delta E V_n^N(\mathbf A_{t}).
\end{equation}
The day value depends on both time-$t$ productivity and on labor allocations decided during the night period, before $A_t$ was known.

Note that choice $L_{njt}$ does not affect future state $A_t$.

The first-order-condition for the night Bellman is
\[
\frac 
	{\partial E_{A} V_n^D(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\partial L_{njt}^*}
	= 0
\]
for all $j$. To determine the partial derivative of the day value, we can use the envelope theorem,
\[
\frac 
	{\partial V_n^D(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\partial L_{njt}^*}
	= 
	\frac \varrho {L_{nt}}
	\left(
				\frac {L_{njt}} {L_{nt}}
				- 
				\frac {L_{njt}^*} {L_{nt}}
	\right).
\]
Equating this with zero in expectation,
\[
L_{njt}^* = E_{A} L_{njt}.
\]
Using the first-order-condition for the day period,
\[
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)}
	{\sum_k w_{nkt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)L_{nkt}/L_{nt}}
\equiv 
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)}
= \lambda+
	\varrho 
	\left(
				\frac {L_{njt}} {L_{nt}}
				- 
				\frac {L_{njt}^*} {L_{nt}}
	\right).
\]
In sectors where the wage rate is higher than the weighted-average wage of the economy, the optimal allocation of labor is higher than the ex-ante expected allocation.

Because in expectation the last term on the right-hand-side of this equation is zero, the $L^*$ has to be such that
\[
E_{A}\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)}
= \lambda
\]
constant. There can be no expected deviations in the (ratio) relative wage across sectors. This is exactly the same condition we were using before,
\[
E_{A}
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)L_{njt}^*}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}, \mathbf L_{t}^*)L_{nt}}
= 
\frac 
	{L_{njt}^*}
	{L_{nt}}.
\]
We can use our existing algorithm to calculate $L^*$, and then $L$ is going to be a random variable with a simple linear combination of $L^*$ and the relative wage.
\end{document}

>>>>

\subsection{Time series}
Given the autoregressive nature of productivity, the vector of $A_{jnt}$s is a state variable. The problem is drastically simplified by the fact that no decision variable affects the state. The representative agent solves
% FIXME: we have already used varrho
\[
\max_{\{L_{njt}, L_{njt}^*\}}
E
\sum_{t=0}^{\infty} 
	\delta^{t} 
	\left[
	\ln \frac 
		{\sum_j w_{njt}(\mathbf A_{t}, \mathbf L_{t})L_{njt}}
		{P_{nt}(\mathbf A_{t}, \mathbf L_{t})}
	-\frac \varrho 2
		\sum_{j=1}^J
		\left(
			\frac {L_{njt}} {L_{nt}}
			- 
			\frac {L_{njt}^*} {L_{nt}}
		\right)^2
	\right]
\]
subject to $\sum_j L_{njt}=\sum_j L_{njt}^*=L_{nt}$ and $w$ and $P$ being determined in equilibrium. The starred labor allocation is determined before shock realizations are known.

The corresponding Bellman equation is
\begin{equation}
	V_n^N(\mathbf A_{t-1}) = 
	\max_{L_{njt}^*}
	E_{A} V_n^D(\mathbf A_{t}, \mathbf L_{t}^*)
\end{equation}
for the ``night'' period before productivity shocks have been realized. For the ``day'' period,
\begin{equation}
	V_n^D(\mathbf A_{t}, \mathbf L_{t}^*) = 
	\max_{\{L_{njt}\}}
		\ln \frac 
			{\sum_j w_{njt}(\mathbf A_{t}, \mathbf L_{t})L_{njt}}
			{P_{nt}(\mathbf A_{t}, \mathbf L_{t})}
		-\frac \varrho 2
			\sum_{j=1}^J
			\left(
				\frac {L_{njt}} {L_{nt}}
				- 
				\frac {L_{njt}^*} {L_{nt}}
			\right)^2
	+ \delta E V_n^N(\mathbf A_{t}).
\end{equation}
The day value depends on both time-$t$ productivity and on labor allocations decided during the night period, before $A_t$ was known.

Note that choice $L_{njt}$ does not affect future state $A_t$. Also, wages and prices do not depend on $L^*$ directly, only through $L$.

The first-order-condition for the night Bellman is
\[
\frac 
	{\partial E_{A} V_n^D(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\partial L_{njt}^*}
	= 0
\]
for all $j$. To determine the partial derivative of the day value, we can use the envelope theorem,
\[
\frac 
	{\partial V_n^D(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\partial L_{njt}^*}
	= 
	\frac \varrho {L_{nt}}
	\left(
				\frac {L_{njt}} {L_{nt}}
				- 
				\frac {L_{njt}^*} {L_{nt}}
	\right).
\]
Equating this with zero in expectation,
\[
L_{njt}^* = E_{A} L_{njt}.
\]
Or, in other words,
\[
L_{njt} = L_{njt}^* + \varepsilon_{njt}(\mathbf A_t),
\]
where $\varepsilon_{njt}$ is the equilibrium deviation from labor, with zero mean conditional on past productivity.

Using the first-order-condition for the day period,
\[
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{\sum_k w_{nkt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)L_{nkt}/L_{nt}}
\equiv 
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
= \lambda L_{nt}+
		\varrho 
				\frac {\varepsilon_{njt}(\mathbf A_t)} {L_{nt}}
.
\]
In sectors where the wage rate is higher than the weighted-average wage of the economy, the optimal allocation of labor is higher than the ex-ante expected allocation.

The challenge is that both sides depend on $\varepsilon_t$, the labor response to productivity innovations.

Because in expectation the last term on the right-hand-side of this equation is zero, the $L^*$ has to be such that
\[
E_{A}\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
= \lambda L_{nt} \equiv 1
\]
constant. There can be no expected deviations in the (ratio) relative wage across sectors.

Substituting in $\lambda = 1/L_{nt}$,
\begin{equation}{eq:varepsilon}
\frac {\varepsilon_{njt}} {L_{nt}}
=
\frac 1\varrho
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)-w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
.
\end{equation}
This can be solved for $\varepsilon$ implicitly, as the RHS is (likely) decreasing in $\varepsilon_{njt}$. We then have $\varepsilon_{njt}(\mathbf A_t, \mathbf L_t^*)$ and $\tilde w_{njt}(\mathbf A_t, \mathbf L_t^*)$. Note that $\tilde w()$ is a different function from $w()$, because it already captures the endogenous labor response. They are only identical when $\varrho=\infty$. Otherwise, the distribution of $\tilde w$ depends on adjustment costs.

The night labor share is a fixed point of
\[
E_{A}\frac 
	{\tilde w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*)}
	{\tilde w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*)}
= 1.
\]
Can we calculate the mean wage share without solving for the entire distribution of $\tilde w$? Totally differentiating equation \eqref{eq:varepsilon},
\begin{equation}{eq:varepsilon'}
\frac {d\varepsilon_{njt}} {L_{nt}}
=
\frac 1\varrho
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)-w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
.
\end{equation}
\end{document}
